### 2018-06-06 Physics Notes/Paper 2/Paper 2.tex
Physics 2 notes

@@

\end{document}

>>>>

\paragraph{Deflection}

@@

For this, use coulomb's equation:
$$F=\frac{1}{4\pi\epsilon_0}\frac{Q_1Q_2}{r^2}$$
\paragraph{Derivation of \large$\mathlarger{R=r_0A^{\frac{1}{3}}}$}
$r_0$ is a constant $1.05fm$\\
$R$ is the radius of the nucleus\\
$A$ is mass number (not activity)\\
Plotting a graph of $\ln R$ against $\ln A$ gives a straight line with gradient $\frac{1}{3}$ and y intercept $\ln r_0$, this gives the equation
$$\ln R=\frac{1}{3}\ln A+\ln r_0$$
By combining the logs, then doing e to both sides, it can be shown that
$$R=r_0A^{\frac{1}{3}}$$
\paragraph{Nuclear density}
$$V=\frac{4}{3}\pi R^3=\frac{4}{3}\pi(r_0 A^{\frac{1}{3}})^3=\frac{4}{3}\pi r_0^3A$$
$$\textrm{Density}=\frac{\textrm{Mass}}{\textrm{Volume}}=\frac{Au}{\frac{4}{3}\pi r_0^3 A}=\frac{u}{\frac{4}{3}\pi r_0^3}=\frac{1.661\times 10^{27}}{\frac{4}{3}\pi(1.05\times10^{-15})^3}=3.4\times10^{17}kgm^{-3}$$
\subsubsection{Mass and energy}
\paragraph{Mass difference and binding energy}
The \textbf{binding energy} of the nucleus is the work that must be done to separate a nucleus into its constituent neutrons and protons.\\
\subparagraph{Example Calculation}

Nucleus of an isotope $^A_ZX$ has:\\
Z Protons\\
A-Z Neutrons\\
Mass=$M_{NUC}$
$$\textrm{Mass defect:} \ \Delta m=Zm_p+(A-Z)m_n-M_{NUC}$$
Where:\\
$m_p=\textrm{Mass of proton} \quad m_n=\textrm{Mass of neutron}$
\\
\\
\textit{The polonium isotope $^{210}_{84}Po$ emits $\alpha$ particles and decays to form the stable isotope of lead $^{206}_{82}Pb$. Write down an equation to represent this process and calculate the energy released when a $^{210}_{84}Po$ nucleus emits an $\alpha$ particle.\\
Mass of a $^{210}_{84}Po$ Nucleus=$209.93667u$\\
Mass of a $^{206}_{82}Pb$ Nucleus=$205.92936u$\\
Mass of a $\alpha$ particle=$4.00150u$\\
$1u=931.3MeV$}\\
$$^{206}_{82}Pb\rightarrow \ ^4_2\alpha+^{206}_{82}Pb+\textrm{Energy released}$$
$$\textrm{Mass difference}=\textrm{Total initial mass-total final mass}$$
$$\textrm{Mass difference}=209.93667-(205.92936+4.00150)$$
$$\textrm{Mass difference}=5.81\times10^{-3}u$$
$$\textrm{Energy released:}\ Q=\textrm{Mass difference in u}\times931.3=5.41MeV$$ 
\subparagraph{Binding energy}
$$E=\Delta mc^2$$
\paragraph{Fission and fusion}
\includegraphics[width=10cm]{Binding_Energy.jpg}\\
In \textbf{nuclear fission} a large unstable nucleus splits into two more stable fragments,  binding energy per nucleon increases.\\
In \textbf{nuclear fusion} two small nuclei fuse to form a larger nucleus with increased binding energy per nucleon\\
Note that the gradient for fusion is much larger than for fusion, meaning that more energy is released in a fusion reaction than a fission reaction.
\subsubsection{Induced fission}
\textbf{Critical mass} - The minimum mass of the fissile isotope in a nuclear reactor necessary to produce a chain reaction\\
\\
The process:\\
\includegraphics[width=6cm]{chain_reaction.png}\\
\\
Components of a nuclear reactor:
\begin{itemize}
\item Moderator - Slows the neutrons down so they can go on to produce further fission
\item Control rods - Rods made of a neutron absorbing substance that are move in or out of the core of a nuclear reactor to control the rate of fission events in the reactor
\item Coolant - A fluid used to prevent the reactor becoming dangerously hot
\end{itemize}
Neutrons undergo elastic collisions with the particles of the moderator, this causes the kinetic energy to transfer from the neutrons to the particles of the moderator, causing the neutrons to slow down.
\paragraph{Factors affecting the choice of materials for the moderator, control rods and coolant}
Moderator:
\begin{itemize}
\item Available in large amounts
\item Small/light nuclei
\item Not become radioactive when bombarded by neutrons
\item High meting point if a solid is used
\end{itemize} $ $\\
Control rods:
\begin{itemize}
\item Absorb neutrons
\item Have a high melting point
\end{itemize} $ $\\
Coolant:
\begin{itemize}
\item High specific heat capacity
\item Non corrosive
\item Stable at high temperatures
\end{itemize}
\subsubsection{Safety aspects}
\begin{itemize}
\item Thick steel reactor core to withstand high pressure/temperature and to absorb $\beta$ radiation and some $\gamma$ radiation and neutrons
\item Building has thick concrete walls to absorb neutrons and $\gamma$ radiation
\item Control rods to stop the reactions in an emergency
\item Fuel rods removed by remote handling devices
\end{itemize}
\end{document}
